### 2013-07-29 Tex/chapters/model.tex
Spacing nach i.e. und e.g. korrigiert
Displacement Beschreibung leicht verbessert
Bildverweise präzisiert
Erwartungen etwas ausführlicher formuliert

    deviation \(\sigma\), i.e. the \(x\) and \(y\) coordinates of the

@@

    \[x \to x + \Delta x\]
    \[y \to y + \Delta y\]

@@

    The boundary is periodic e.g. nodes near the right edge can be

@@

    for which exists an analytic solution \cite{Onsager1944}. And for
    \(\sigma \gtrsim 1\) the nodes are distributed randomly. This case
    is already studied for constant \(J\) on the Delaunay triangulation
    in \cite{Janke1994}.\\

@@

    The Gabriel graph (GG) \cite{Gabriel1969} is a subgraph of the
    Delaunay triangulation. Two nodes \(i\) and \(j\) with distance
    \(d_{ij}\) are connected with an edge, if a circle with its
    center on half way between \(i\) and \(j\) and radius
    \(r = \frac d 2\) contains no other nodes. This area will be
    called \emph{lune} in the following. See also Figure
    \ref{fig:lunes}\subref{sfig:lunes:def}.\\
    The Relative Neighborhood graph (RNG) \cite{Toussaint1980} is a
    subgraph of the GG. Two nodes \(i\) and \(j\) with
    distance \(d_{ij}\) are connected, if no other node is in the
    \emph{lune}. The lune is defined as the intersection of two
    circles with radius \(r = d\) and centers on \(i\) and \(j\).
    See also Figure \ref{fig:lunes}\subref{sfig:lunes:def}.
    \begin{figure}[htbp]
        \centering
        \subfigure[Definition of the Lunes][]{
            \label{sfig:lunes:def}
            \input{images/GG_RNG_def}
        }
        \subfigure[RNG example][]{
            \label{sfig:lunes:rng}
            \includegraphics[width=0.3\textwidth]{images/RNG/L12S03.pdf}
        }
        \subfigure[GG example][]{
            \label{sfig:lunes:gg}
            \includegraphics[width=0.3\textwidth]{images/GG/L12S03.pdf}
        }
        \caption[Gabriel - and Relative Neighborhood Graph]
        {
            \subref{sfig:lunes:def} Lunes of RNG (hatched region) and
                GG (cross hatched region)
            \subref{sfig:lunes:rng} Example of a RNG on periodic
                boundary conditions. Periodic nodes are dashed.
            \subref{sfig:lunes:gg} Example of a GG on
                periodic boundary conditions. Periodic nodes are dashed.
        }
        \label{fig:lunes}
    \end{figure}\\
    To construct these graphs the simple way is to test for each
    pair of nodes if any other node lies in
    the lune of the pair. That is of complexity \(O (N^3)\), because
    there are \(N(N-1)\) pairs and for each \(N-2\) nodes to test. So
    the product is of order \(O(N^3)\)\\
    To reduce the complexity one can first create a Delaunay
    Triangulation in complexity \(O (N \log N)\)
    \cite{RNGCell} and test the criterion for each edge, because
    the Delaunay triangulation is a supergraph of both. But the
    implementation of a Delaunay triangulation algorithm is not trivial
    and the generation of the graphs is not time critical in the scope
    of this bachelor thesis.\\
    So a trade off is to use basically the simple method but only test
    the criterion for nodes which are near to the lune and abort if
    one node inside the lune is found. To determine which nodes are
    near the lune one can subdivide the area in \emph{cells} and save
    for each cell a list with nodes lying inside it like presented in
    \cite{RNGCell}.
    Now it is just necessary to test the nodes in the cells which
    resemble a rectangular bounding box of the lune. Most pairs will be
    far away from each other and the cells in the middle of the bounding
    box are completely inside the lune so that only one node has to be
    tested to discard an edge between them. Connected nodes are near to
    each other so that only very few cells have to be tested.\\
    Indeed this method reduced the time needed to construct a RNG with
    \(N=32^2\) and \(N=64^2\) by a factor of
    over \(15\) respectively \(40\). Though the complexity is still of
    order \(O(N^2)\) in the best case, because for every pair at least
    one check has to be performed.

>>>>

    deviation \(\sigma\), i.e.\ the \(x\) and \(y\) coordinates of the

@@

@@

    The boundary is periodic e.g.\ nodes near the right edge can be

@@

    for which exists an analytic solution \cite{Onsager1944} for the
    critical temperature
        T_c = 2J/\ln(1+\sqrt 2) = 2.2691...
        \label{eq:exactTc}
    and the critical exponents \(\alpha = 0\), \(\beta = \frac{1}{8}\)
    and \(\gamma = -\frac{7}{4}\).\\
    The case for randomly distributed sites \(\sigma \gtrsim 1\) is
    already studied on a Delaunay triangulation. \cite{Janke1994} examines
    this for constant \(J\) and \cite{Lima2000} for \(J_{ij} = e^{-\alpha d_{ij}}\).
    Both articles conclude, that this model lies within the universality
    class of the standard Ising model, i.e.\ has the same critical
    exponents.

@@

    The Gabriel graph (GG) \cite{Gabriel1969} is a subgraph of the
    Delaunay triangulation. Two nodes \(i\) and \(j\) with distance
    \(d_{ij}\) are connected with an edge, if a circle with its
    center on half way between \(i\) and \(j\) and radius
    \(r = \frac d 2\) contains no other nodes. This area will be
    called \emph{lune} in the following. See also the cross hatched region
    from fig. \ref{fig:lunes}\subref{sfig:lunes:def}.\\
    The Relative Neighborhood graph (RNG) \cite{Toussaint1980} is a
    subgraph of the GG. Two nodes \(i\) and \(j\) with
    distance \(d_{ij}\) are connected, if no other node is in the
    \emph{lune}. The lune is defined as the intersection of two
    circles with radius \(r = d\) and centers on \(i\) and \(j\).
    See also the hatched region in fig. \ref{fig:lunes}\subref{sfig:lunes:def}.
    \begin{figure}[htbp]
        \centering
        \subfigure[Definition of the Lunes][]{
            \label{sfig:lunes:def}
            \input{images/GG_RNG_def}
        }
        \subfigure[RNG example][]{
            \label{sfig:lunes:rng}
            \includegraphics[width=0.3\textwidth]{images/RNG/L12S03.pdf}
        }
        \subfigure[GG example][]{
            \label{sfig:lunes:gg}
            \includegraphics[width=0.3\textwidth]{images/GG/L12S03.pdf}
        }
        \caption[Gabriel - and Relative Neighborhood Graph]
        {
            \subref{sfig:lunes:def} Lunes of RNG (hatched region) and
                GG (cross hatched region)
            \subref{sfig:lunes:rng} Example of a RNG on periodic
                boundary conditions. Periodic nodes are dashed.
            \subref{sfig:lunes:gg} Example of a GG on
                periodic boundary conditions. Periodic nodes are dashed.
        }
        \label{fig:lunes}
    \end{figure}\\
    To construct these graphs the simple way is to test for each
    pair of nodes if any other node lies in
    the lune of the pair. That is of complexity \(O (N^3)\), because
    there are \(N(N-1)\) pairs and for each \(N-2\) nodes to test. So
    the product is of order \(O(N^3)\)\\
    To reduce the complexity one can first create a Delaunay
    Triangulation in complexity \(O (N \log N)\)
    \cite{RNGCell} and test the criterion for each edge, because
    the Delaunay triangulation is a supergraph of both. But the
    implementation of a Delaunay triangulation algorithm is not trivial
    and the generation of the graphs is not time critical in the scope
    of this bachelor thesis.\\
    So a trade off is to use basically the simple method but only test
    the criterion for nodes which are near to the lune and abort if
    one node inside the lune is found. To determine which nodes are
    near the lune one can subdivide the area in \emph{cells} and save
    for each cell a list with nodes lying inside it like presented in
    \cite{RNGCell}.
    Now it is just necessary to test the nodes in the cells which
    resemble a rectangular bounding box of the lune. Most pairs will be
    far away from each other and the cells in the middle of the bounding
    box are completely inside the lune so that only one node has to be
    tested to discard an edge between them. Connected nodes are near to
    each other so that only very few cells have to be tested.\\
    Indeed this method reduced the time needed to construct a RNG with
    \(N=32^2\) and \(N=64^2\) by a factor of
    over \(15\) respectively \(40\). Though the complexity is still of
    order \(O(N^2)\) in the best case, because for every pair at least
    one check has to be performed.
